Multi-period and Multi-product Batch Processing Time Maximization Problem

### 2023-08-02 multi-period/manuscript.tex
elaborating solution method for a multi-period problem

\emph{Example:} A certain machine must process a batch containing 2 different products: A and B. The production rate of A is 60 g/min while the production rate of B is 40 g/min. The factory has free stock for a maximum of 3000 g of any product, and, according to the company's inventory policy, a maximum of 3000 g of product A and 2000 g of product B can be stocked at the factory. For the first day, there is a demand for 1000 g of product A and 500 g of product B, and for the second day, there is a demand for 500 g of product A and 300 g of product B. The factory has an outlet that has free space in stock of 1000 g, which can receive a maximum of 600 g of each product.There is a previous production planning for 1000 units of product A for the first day and 800 units of product B for the second day. A maximum time of 100 minutes of this machine can be allocated for processing this batch. What is the maximum possible time for processing this batch ?  \\

@@

\begin{equation}
P_i + UI_{i1} - D_{i1} - O_{i1} = \textrm{UI}_{i0} - \textrm{PO}_{i1}
\end{equation}

\begin{equation}
UI_{id} - UI_{i(d-1)} - D_{id} - O_{id} = - \textrm{PO}_{id} \quad \forall i, \forall (d>1)
\end{equation}

\begin{equation}
I_{1} + \sum_{i}\{P_i -  D_{i1} - O_{i1}\} = \textrm{I}_{0} 
\end{equation}

\begin{equation}
I_{d} - I_{d-1} + \sum_{i}\{P_i -  D_{id} - O_{id}\} = 0 \quad \forall (d>1)
\end{equation}

@@

\begin{equation}
D_{i1} \leq \textrm{UD}_{i1} \quad \forall i
\end{equation}

@@

\section{Analytical solution}
\label{sec:analyticalSol}

It is probably possible to extend this thinking to the solution of a class of linear optimization problems. However, we will work on this concept in an upcoming paper. \\

@@

$T' \leq \lfloor{1000 \textrm{ g} / (60 \textrm{ g}/\textrm{min})}\rfloor$ \\

$T' \leq 16 \textrm{ min}$ \quad for A \\

$T' \leq \lfloor{500 \textrm{ g} / (40 \textrm{ g}/\textrm{min})}\rfloor$ \\

$T' \leq 12 \textrm{ min}$ \quad for B \\

so: \\

$T' = 12 \textrm{ min}$ \\

@@

$S_A = 1000 \textrm{ g} - 12 \textrm{ min} * 60 \textrm{ g}/\textrm{min} = 280 \textrm{ g}$ \\

$S_B = 500 \textrm{ g} - 12 \textrm{ min} * 40 \textrm{ g}/\textrm{min} = 20 \textrm{ g}$ \\

@@

$T'' \leq \lfloor{(3000 \textrm{ g} + 600 \textrm{ g} + 280 \textrm{ g}) / (60 \textrm{ g}/\textrm{min})}\rfloor$ \\

$T'' \leq 64 \textrm{ min}$ \quad for A \\

$T'' \leq \lfloor{(2000 \textrm{ g} + 600 \textrm{ g} + 20 \textrm{ g}) / (40 \textrm{ g}/\textrm{min})}\rfloor$ \\

$T'' \leq 65 \textrm{ min}$ \quad for B \\

$T'' \leq \lfloor{(1000 \textrm{ g} + 3000 \textrm{ g} + 300 \textrm{ g}) / (100 \textrm{ g}/\textrm{min})}\rfloor$ \\

$T'' \leq 43 \textrm{ min}$ \quad for A and B \\

@@

$T'' = 43 \textrm{ min}$ \\

@@

$T = 55 \textrm{ min}$ \\

As $T<\textrm{Z}$, then $T = 55 \textrm{ min}$ will certainly be the optimal solution. \\

@@

$P_A = 3300 \textrm{ g}$ e $P_B=2200 \textrm{ g}$ \\

$D_A = 1000 \textrm{ g}$ e $D_B = 500 \textrm{ g}$ \\

$E_A = 2300 \textrm{ g}$ e $E_B = 1700 \textrm{ g}$ \\

Note that it is not important to know the values of $O_i$ and $I_i$, $\forall i$, however, after determining a solution, these values can be found by taking a solution from the undetermined system of inequalities:

\begin{equation}
O_i + I_i = E_i \quad \forall i
\end{equation}

\begin{equation}
O_i \leq \textrm{UO}_i \quad \forall i
\end{equation}

\begin{equation}
\sum_i{O_i} \leq \textrm{O}
\end{equation}

\begin{equation}
I_i \leq \textrm{UI}_i \quad \forall i
\end{equation}

\begin{equation}
\sum_i{I_i} \leq \textrm{I}
\end{equation}

\begin{equation}
O_i, I_i \in  \mathbb{Z}^+ \quad \forall i
\end{equation}

@@

We are enormously grateful to Coordenação de Aperfeiçoamento de Pessoal de Nível Superior (CAPES) and to Conselho Nacional de Desenvolvimento Científico e Tecnológico (CNPq) for the financial support provided to our projects. We also thank LINDO systems team for the LINGO software license, without which this work would not have been possible and the to the owner of the company in the plastics sector, who allowed us to learn about his company's production process. Finally, we would like to thank Pró-reitoria de Extensão e Cultura da UFPE (PROExC) and the Research Director of Propesqi (Pró-reitoria de Pesquisa e Inovação da UFPE) for their support and recognition of our work, and dear Professors Antônio José da Silva Neto and João Flávio Vieira de Vasconcellos from IPRJ/UERJ, who contributed significantly to the formation of essential skills for the development of our projects. We also thank my co-worker Marcos Luiz Henrique, for having helped by evaluating the mathematical model and solution method proposed in this paper.

>>>>

\emph{Example:} A certain machine must process a batch containing 2 different products: A and B. The production rate of A is 60 g/min while the production rate of B is 40 g/min. The factory has free stock for a maximum of 3000 g of any product, and, according to the company's inventory policy, a maximum of 3000 g of product A and 2000 g of product B can be stocked at the factory. For the first day, there is a demand for 1000 g of product A and 500 g of product B, and for the second day, there is a demand for 500 g of product A and 300 g of product B. The factory has an outlet that has free space in stock of 1000 g, which can receive a maximum of 600 g of each product. There is a previous production planning for 1000 units of product A for the first day and 1500 units of product B for the second day. A maximum time of 100 minutes of this machine can be allocated for processing this batch. What is the maximum possible time for processing this batch ?  \\

@@

\begin{equation}
\label{eq:mmeq02}
P_i + UI_{i1} - D_{i1} - O_{i1} = \textrm{UI}_{i0} - \textrm{PO}_{i1}
\end{equation}

\begin{equation}
\label{eq:mmeq03}
UI_{id} - UI_{i(d-1)} - D_{id} - O_{id} = - \textrm{PO}_{id} \quad \forall i, \forall (d>1)
\end{equation}

\begin{equation}
I_{1} + \sum_{i}\{P_i -  D_{i1} - O_{i1}\} = \textrm{I}_{0} - \sum_{i}{\textrm{PO}_{i1}}
\end{equation}

\begin{equation}
I_{d} - I_{d-1} + \sum_{i}\{P_i -  D_{id} - O_{id}\} = - \sum_{i}{\textrm{PO}_{id}} \quad \forall (d>1)
\end{equation}

@@

\begin{equation}
\label{eq:mmeq09}
D_{i1} \leq \textrm{UD}_{i1} \quad \forall i
\end{equation}

@@

\section{Exact optimizatin method}
\label{sec:method}

Uma solução ótima para este problema pode ser construída através de duas etapas, conforme explicado a seguir: \\

\emph{etapa 1: construção do MPBPTMP} \\

passo 1: Desconsidere a produção a ser planejada:

\begin{equation}
\label{eq:sm01}
P_i = 0, \forall i
\end{equation}

passo 2: Calcule os valores $D_{id}^+$, $D_{id}^-, \forall i, d$, usando as equações em (\ref{eq:sm02}) e (\ref{eq:sm03}):

\begin{equation}
\label{eq:sm02}
D_{id}^+ = \sum_{k=1}^{d}{\textrm{UD}_{ik}} - \sum_{k=1}^{d-1}{D_{ik}}  \quad \forall i, \forall d>1
\end{equation}

\begin{equation}
\label{eq:sm03}
D_{id}^- = \sum_{k=2}^{d}{\textrm{UD}_{ik}} - \sum_{k=1}^{d-1}{D_{ik}}  \quad \forall i, \forall d>1
\end{equation}

Calcule $D_{i1}$ usando as equações em (\ref{eq:sm04}):

\begin{equation}
\label{eq:sm04}
D_{i1} = \min \{\textrm{UD}_{i1}, \textrm{PO}_{i1}\} \quad \forall i
\end{equation}

queremos encontrar $D_{id}$, tal que a equação em (\ref{eq:sm04}) seja respeitada:

\begin{equation}
\label{eq:sm04}
D_{id} = \min {x \in \{D_{id}^-, D_{id}^+\}} \textrm{ tal que } UI_{id} \geq 0
\end{equation}

Contudo, inicialmente iremos adotar:

\begin{equation}
\label{eq:sm04}
D_{id} = D_{id}^- \quad \forall i, \forall d>1
\end{equation}

e calcule os valores de $D_{id}$ e $O_{id}, \forall i, d$.

Inicialmente desconsidere a produção a ser planejada fazendo $P_i = 0, \forall i$, e

Tais valores podem ser calculados da seguinte forma:

- distribua a produção previamente planejada para atender à demanda seguindo as restrições em (9) e para as duas equações derivadas de (10):

as restrições em (14), (15) e (16) foram criadas porque queremos permitir que a demanda referênte ao dia atual seja atendida pela produção planejada para o dia atual e não por produções futuras.

- distribua o restante da produção previamente planejada, se houver, para os outlests respeitando as restrições em (7) e (8).

- faça $D_{id} = D_{id}^- \quad \forall i$;

- calcule os valores para $UI_{id}, \forall i, d$, usando as equações em (3) e (4).

- calcule os valores para $I_{i}, \forall i$, usando as equações em (5) e (6).

- recalcule o valor de $D_{id}$, usando:

\begin{equation}
D_{id} = D_{id}^- - \max\{0, UI_{id}\} \quad \forall i, d>0
\end{equation}

E então recalcule $UI_{id}, \forall i, (d>1)$, usando as equações em (4).

Então, com base nos valores claculado, construímo o problema MPBPTM, usando as equações:

\begin{equation}
\textrm{UD}_{i} = \textrm{UD}_{i1} - D_{i1}, \quad \forall i
\end{equation}

\begin{equation}
\textrm{I} = \min_{d} \{I_{d}\}
\end{equation}

\begin{equation}
\textrm{UI}_{i} = \min_{d} \{UI_{id}\}, \quad \forall{i}
\end{equation}

\begin{equation}
\textrm{O} = \textrm{O} - \sum_{i}{\sum_{d}{O_{id}}}
\end{equation}

\begin{equation}
\textrm{UO}_{i} = \textrm{UO}_{i} - \sum_{d}{O_{id}}, \quad \forall i
\end{equation} \\

Sendo $\textrm{p}_{i}$ e $\textrm{Z}$ definidos de acordo com o problema apresentado.

E resolvemos esse novo problema usando o método de partição do tempo proposto por Fraga (Fraga et al., 2023).\\

@@

Como planejamento prévio de produção, temos:

$\textrm{P}_{11} = 1000$,
$\textrm{P}_{12} = 0$,
$\textrm{P}_{21} = 0$,
$\textrm{P}_{22} = 1500$. \\

Como demanda, temos:

$\textrm{UD}_{11} = 1000$,
$\textrm{UD}_{12} = 500$,
$\textrm{UD}_{21} = 500$,
$\textrm{UD}_{22} = 300$. \\

Distribuindo a produção para atender à demanda de acordo com as restrições em (9) e (10), temos: \\

$D_{11} = 1000$,
$D_{12} = 0$,
$D_{21} = 0$,
$D_{22} = 300$. \\

Não há nenhuma sobra relativa a produção de A. E há uma sobra de produção de 1200 unidades de B no segundo dia. \\

Como estoque livre nos outlets, temos:

$\textrm{UO}_{1} = 600$;
$\textrm{UO}_{2} = 600$;
$\textrm{O} = 1000$; \\

Assim, distribuindo a sobra da produção para os outlets de acordo com as restrições em (7) e (8), temos:

$O_{11} = 0$,
$O_{12} = 0$,
$O_{21} = 0$,
$O_{22} = 600$. \\

Com relação ao estoque livre em fábrica, temos:

$\textrm{UI}_{10} = 3000$,
$\textrm{UI}_{20} = 2000$,
$\textrm{I}_{0} = 3000$. \\

Portanto, calculando os valores disponíveis para produção de acordo com as equações em (3), (4) e (5), temos:

$UI_{11} = 3000 - 1000 + 1000 + 0 =  3000$,
$UI_{12} = 3000 - 0 + 0 + 0 =  3000$,
$UI_{21} = 2000 - 0 + 0 + 0 =  2000$,
$UI_{22} = 2000 - 1500 + 300 + 600 =  1400$. \\

$I_{1} = 3000 - 1000 + 1000 + 0 = 3000$,
$I_{2} = 3000 - 1500 + 300 + 600 = 2400$. \\

Portanto temos agora as informações referentes aos limites estocagem em fábrica diário para cada produto assim como o limite de estocagem em fábrica diário para todos os produtos. \\

Definindo o problema MPBPTM, temos: \\

$\textrm{UD}_{1} = 1000 - 1000 = 0$,
$\textrm{UD}_{2} = 0$ . \\

$\textrm{I} = \min_{d} \{3000, 2400\} = 2400$. \\

$ \textrm{UI}_{1} = \min_{d} \{3000, 3000\} = 3000 $,
$ \textrm{UI}_{2} = \min_{d} \{2000, 1400\} = 1400 $. \\

$\textrm{O} = 1000 - 600 = 400.$ \\

$\textrm{UO}_{1} = 600 - 0 = 600$, 
$\textrm{UO}_{2} = 600 - 600 = 0$. \\

$\textrm{p}_{1} = 60$,
$\textrm{p}_{2} = 40$.  \\

$\textrm{Z} = 100$.  \\

Então, aplicando o método de partição do tempo proposto pro Fraga, temos: \\

$T' = 0 \textrm{ min}$ \\

@@

$S_A = 0 \textrm{ g}$ \\

$S_B = 0 \textrm{ g}$ \\

@@

$T'' \leq \lfloor{(3000 \textrm{ g} + 600 \textrm{ g} + 0 \textrm{ g}) / (60 \textrm{ g}/\textrm{min})}\rfloor$ \\

$T'' \leq 60 \textrm{ min}$ \quad for A \\

$T'' \leq \lfloor{(1400 \textrm{ g} + 0 \textrm{ g} + 0 \textrm{ g}) / (40 \textrm{ g}/\textrm{min})}\rfloor$ \\

$T'' \leq 35 \textrm{ min}$ \quad for B \\

$T'' \leq \lfloor{(400 \textrm{ g} + 2400 \textrm{ g} + 0 \textrm{ g}) / (100 \textrm{ g}/\textrm{min})}\rfloor$ \\

$T'' \leq 28 \textrm{ min}$ \quad for A and B \\

@@

$T'' = 28 \textrm{ min}$ \\

@@

$T = 28 \textrm{ min}$ \\

As $T<\textrm{Z}$, then $T = 28 \textrm{ min}$ will certainly be the optimal solution. \\

@@

$P_A = 1,680  \textrm{ g}$ e $P_B=1,120 \textrm{ g}$ \\

$D_A = 0 \textrm{ g}$ e $D_B = 0 \textrm{ g}$ \\

$E_A = 1,680 \textrm{ g}$ e $E_B = 1,120 \textrm{ g}$ \\

@@

We are enormously grateful to Coordenação de Aperfeiçoamento de Pessoal de Nível Superior (CAPES) and to Conselho Nacional de Desenvolvimento Científico e Tecnológico (CNPq) for the financial support provided to our projects. We also thank LINDO systems team for the LINGO software license, without which this work would not have been possible and the to the owner of the company in the plastics sector, who allowed us to learn about his company's production process. Finally, we would like to thank Pró-reitoria de Extensão e Cultura da UFPE (PROExC) and the Research Director of Propesqi (Pró-reitoria de Pesquisa e Inovação da UFPE) for their support and recognition of our work, and dear Professors Antônio José da Silva Neto and João Flávio Vieira de Vasconcellos from IPRJ/UERJ, who contributed significantly to the formation of essential skills for the development of our projects.
